Lattice Boltzmann simulations of magnetically responsive\\ ellipsoidal particle stabilized emulsions

### 2025-03-24 chapters/lit_review.tex
Changes 1. Updated lit review text for flow and clarity. Still have work to do here

\section{Emulsion stability and microstructures}

When two partially miscible fluids are mixed, they tend to separate into distinct phases to minimize the thermodynamic penalty associated with interfacial formation. 
To inhibit this phase separation, small-molecule surfactants are conventionally employed. A well-known example of such a material is soap, which facilitates the 
emulsification of dirt into droplets suspended in water through mechanical agitation. Surfactants function by reducing the surface tension between the dispersed 
phase (droplets) and the continuous phase (bulk fluid), thereby lowering the interfacial energy penalty associated with phase formation.

The existence of particle-stabilized emulsions has been recognized for over a century, following the independent discoveries by Pickering and Ramsden, who observed dispersed 
oil droplets within a water matrix after vigorous stirring of an oil-water-particle mixture \cite{ramsden_separation_1904, pickering_cxcvi.emulsions_1907}. Unlike surfactants, 
which stabilize emulsions by reducing interfacial tension, particles act as stabilizers by adsorbing at the interface between the dispersed and continuous phases, thereby 
preventing direct contact between them. The Pieranski model is commonly used to determine the adsorption 
energy of a particle at an interface. It is expressed as $ G_{ads} = \sigma A_{rm} (1 - \cos{\theta_c})^2 $ where $G_{ads}$ represents the free energy reduction upon particle 
adsorption, $\sigma$ is the surface tension between the partially miscible fluids, and $ \theta_c $ is the contact angle of the particle. Particles at the interface are 
generally considered irreversibly adsorbed, even at the nanoscale. For particles larger than 100 nm, the adsorption energy is sufficiently high that thermal fluctuations at 
the interface can be considered negligible \cite{cheung_molecular_2011}.

@@

Compared to surfactant-stabilized emulsions, particle-stabilized emulsions exhibit greater resistance to coarsening, leading to renewed interest in their applications. 
The microstructural properties of these emulsions were extensively studied in the early 2000s by Lumsdon, Binks, and others. Their findings indicated that the emulsion 
droplet radius follows the relationship $R_e \propto \frac{\phi_w}{\phi_p}$, where $\phi_w$ and $\phi_p$ represent the volume fractions of water and particles, 
respectively \cite{binks_pickering_2001}. Additionally, they identified several factors influencing the microstructure of Pickering emulsions, including fluid concentration 
and particle wettability. Neutrally wetting particles do not impose a preferential curvature on the interface, whereas non-neutrally wetting particles can lead to the 
formation of bridged droplets, capillary aggregates, and other non-spherical microstructures. A notable example is bijels, which are synthesized in systems containing equal 
volume fractions of immiscible fluids and neutrally wetting particles, leading to the formation of tortuous, co-continuous domains.

Bijels are synthesized in systems containing approximately equal fluid volume fractions and neutrally wetting particles to prevent the imposition of preferential 
curvature at the interface \cite{stratford_colloidal_2005, herzig_bicontinuous_2007, lee_bicontinuous_2010, jansen_bijels_2011, velankar_non-equilibrium_2015}. 
Reeves et al. demonstrated that bijels stabilized with nanoparticles exhibit greater stability than those stabilized with microparticles due to enhanced mechanical 
properties and improved interfacial coverage \cite{reeves_particle-size_2015}. 
Furthermore, research by Jansen, Harting, and Hijnen et al. established that the bijel domain size follows the relationship $ L \propto \frac{1}{\phi_p} $ for both 
rod-like and spherical particles. While the proportionality constant depends on particle geometry, this trend has been observed across various particle shapes 
\cite{hijnen_bijels_2015, madivala_exploiting_2009, gunther_timescales_2014, daware_emulsions_2015, loudet_capillary_2005, cheng_shape-anisotropic_2013}.

% While bijels are a relatively new material class, much work has been conducted to elucidate the underlying mechanisms controlling their stability, microstructure and rheology. Many studies focus on how the particle properties play important roles in tuning the properties of bijels. This section will cover a brief overview and summary of literature on the working principle behind particle surfactants, the effect of size and concentration of particles, anisotropic particles in particle stabilized emulsions, microstructure control in various bijel synthesis techniques, stimuli response in bijels and bijel rheology.

@@

\section{Particle shape anisotropy and synthesis techniques}

Over the past decade, advancements in synthesis techniques have significantly expanded the ability to fabricate particles with anisotropic geometry and surface chemistry. 
The choice of synthesis method depends on the desired particle shape. For example, ellipsoidal particles can be readily produced through the mechanical deformation of 
polymer spheres, while dumbbell-shaped particles can be synthesized using microfluidic devices or emulsion templating. Square-shaped particles, on the other hand, can be 
generated through controlled crystallization \cite{morgan_understanding_2013}. The increasing variety of synthesis methods has reduced geometric constraints when exploring 
potential particle stabilizers for bijels \cite{wu_recent_2016}.

Anisotropic particles exhibit shape-dependent properties due to their ability to induce multipolar interactions by deforming the interface to maintain a mean curvature of 
zero, thereby satisfying the Young-Laplace equation, as illustrated in Figure \ref{fig:anisotropic_particle_interface} \cite{loudet_capillary_2005, cheng_shape-anisotropic_2013}.
This property has been leveraged to facilitate directed assembly and migration through modified capillary forces 
\cite{cavallaro_curvature-driven_2011, read_dimerization_2020, sharifi-mood_curvature_2015}.

When stabilizing bijels, ellipsoidal particles provide greater stability than spherical ones due to their higher cross-sectional area-to-volume ratio, as demonstrated by 
Günther et al. and Hijnen et al. \cite{gunther_timescales_2014, hijnen_bijels_2015}. This enhanced stability arises from additional domain coarsening timescales associated 
with particle reorientation and increased mechanical rigidity, as confirmed by rheological studies \cite{gunther_timescales_2014, daware_emulsions_2015, witt_bijel_2013}. 
Furthermore, particle shape influences both the onset and dynamics of jamming. Studies using graphene plate-like particles have shown that these particles exhibit intrinsic 
elasticity, which affects the conditions under which jamming occurs \cite{imperiali_simple_2014, sun_assembly_2013}.

The orientation of ellipsoidal particles at interfaces is governed by a balance between interfacial capillary forces and external magnetic fields 
\cite{bresme_orientational_2007, davies_assembling_2014}. Theoretical studies and simulations suggest the existence of a critical field strength beyond which particle 
orientation is predominantly dictated by the applied field \cite{bresme_orientational_2007, davies_assembling_2014}. Additionally, both experimental and computational 
studies have shown that steric interactions—modulated by particle orientation and interfacial arrangements—play a significant role in determining the free energy of particle 
assemblies, underscoring the intricate interplay between these forces \cite{morgan_understanding_2013, newton_influence_2014, newton_capillary_2018}.

% The adsorption process is affected by the shape of particles used as they pack differently onto the interface \cite{hijnen_bijels_2015, daware_emulsions_2015,carmack_diverse_2017}. It has also been suggested that adsorption dynamics of ellipsoidal particles at liquid interfaces are driven completely by viscous forces even if the timescales of adsorption are driven by particle properties \cite{Coertjens2017}. Some guiding equations to calculate the interfacial area of an ellipsoidal particle are provided \cite{gunther_timescales_2014, Davies2014}.

The existence of bijels was first identified computationally using a multicomponent lattice boltzmann method which looked at 
the jamming of spherical particles in a binary fluid system with equal volume fraction of each fluid. 
\cite{stratford_colloidal_2005} Once phase separation was initiated, the particles adsorb onto the interface until the interfacial
area matches the cross sectional area of the particles at the interface, when domain coarsening is ceased. The simulation assumed
an equal volume fraction of fluid components and neutrally wetting particles to not impart preferential curvature onto the interface.
Figure \ref{fig:state_diagram_particle_emulsions} demonstrates a schematic derived from experimental results of particle stabilized
emulsion microstructures that form when changing the volume fraction and particle contact angle.

In experiments, bijels were first synthesized using Thermally Induced Phase Separation (TIPS) \cite{herzig_bicontinuous_2007, lee_bicontinuous_2010, bai_dynamics_2015}. 
The process begins by identifying a small-molecule or polymer blend with a critical point, referred to as the casting mixture, and preparing it at a composition 
corresponding to this critical point. The casting mixture is then combined with particle surfactants while remaining in a single-phase state. Once the particles 
are homogeneously distributed, phase separation is induced by adjusting the temperature to bring the system into the two-phase region. By ensuring that the 
composition aligns with the critical point, the casting mixture undergoes spinodal decomposition, preventing nucleation and growth.

TIPS enables batch synthesis of bijels in research laboratories. However, for bijels to become industrially viable, a continuous fabrication process is required—one 
that also allows for the production of various bijel form factors suited to different applications. Solvent Transfer Induced Phase Separation (STrIPS) has been proposed 
as a method to achieve this. In STrIPS, a casting mixture is prepared, consisting of two fluids that form the bijel, a co-solvent, particles, and a surfactant. Phase 
separation is initiated as the co-solvent diffuses out of the bijel. The surfactant is included to maintain bijel stability against Marangoni forces. The casting mixture 
is then extruded into a bath of a non-solvent, which is immiscible with the bijel-forming fluids but miscible with the co-solvent. The diffusion of the co-solvent triggers 
phase separation, leading to bijel formation. The resulting microstructure is influenced by both the surfactant concentration and the flow rate of the casting mixture.

Vapor Induced Phase Separation(VIPS) has also been identified as a technique for continuous fabrication of bijels. \cite{wang_scalable_2020} A quarternary casting mixture
containing particles, solvent and two partially miscible liquids is prepared. The solvent and liquid species are carefully selected to ensure that removal of the
solvent will cause phase separation of the two partially miscible liquids. One system that has been used is the water/hexanediol-diacrylate solvated by ethanol.
Selection of the composition of the system facilitates crossing of the binodal through the critical point, causing phase separation through spinodal decomposition.
This technique has been used to fabricate thin films of bijels blade and spray coated onto substrates.

Instead of relying upon spinodal decomposition to generate the bijel morphology, homogenization uses shear to join phase separating domains, resulting in a 
structure that has the properties of a bijel even if not fabricated from fluids undergoing spinodal decomposition. \cite{huang_bicontinuous_2017, cai_bijels_2017} 
This method is primarily used for nanoparticles under $50$ nm and it has been shown to work for multiple particle geometries. This method relies upon
shear to cause limited coalescence of droplets. As coarsening occurs, particles adsorb on the interface and when the interfacial area matches the area of the
adsorbed particles, the microstructure jams. Tuning of the method can be done through modifying the contact angle of the particles at the interface as well as
the viscosity of the fluids.

% \textcolor{blue}{Description of liquid in liquid printing}

% Enter something about this method. \cite{amirfattahi_fabrication_2024}

\section{Rheological models and shear response of bijels}

@@

@@

\section{Microstructure control in bijels}

@@

\section{Lattice Boltzmann Method}

The Lattice Boltzmann Method (LBM) is an evolution of preceding lattice gas automata techniques, which is a discretization of the Boltzmann equation of motion for 
molecules. This means that unlike traditional CFD techniques such as FDM, VOF or level set methods, LBM is a psuedo-molecule method that tracks the evolution of a 
particle distribution function within grid cells evolved through a discretized Boltzmann equation of motion. Macroscopic variables such as fluid density and 
velocity are recovered from the particle distributions through appropriate moment integration, and the Navier Stokes equation at the incompressible limit can be 
obtained through a Chapman-Enskog expansion of the LBM. The LBM has become an attractive tool for meso-scale CFD simulations due to its ease of algorithm 
implementation, highly parallelizable nature and ease of boundary condition implementation, allowing coupling to other physically relevant systems such as 
particles with varieties of potentials, external fields and deformable bodies.

The particle distribution function described earlier is advected on a pre-constructed lattice stencil, commonly denoted as DnQm where n and m represent the 
number of dimensions and directions in the stencil. Common stencils include the D1Q5, D2Q9 and D3Q19 stencil, all of which recover mass and momentum conservation. 
For energy conservation, a higher order stencil such as D3Q27 is necessary. The D2Q9 and D3Q19 stencils have 8 and 18 populations respectively that include 
connections to nearest and next nearest neighbour points, in addition to a central rest point.

The LBM is composed of a collision and advection step. In the advection step, the populations at each grid point are propagated to adjacent points in accordance 
with the chosen stencil. During the collision step, the particle population distribution is relaxed towards an equilibrium with a collision operator, at a 
specified relaxation rate. The collision operator can have multiple forms based on how many relaxation rates are used although the most common variant is the 
Single Relaxation Time (SRT) collision operator, more commonly known as the Bhatnagar-Gross-Krook (BGK) collision operator. Owing to its stability in low Mach 
and Reynolds numbers and simplicity of implementation, the BGK operator is often used in common particle laden flow and soft matter scenarios. 
\cite{bhatnagar_model_1954} These limitations are present owing to the implicit link between the fluid properties and the relaxation rate necessitating 
relaxation rates above 0.5, and to ensure fluid incompressibility from an equation that intrinsically simulates a compressible fluid. To get over these 
limitations, Two Relaxation Time (TRT) and Multiple Relaxation Time (MRT) operators also exist, expanding the possible application of the LBM to visco-elastic 
flows and implementation of fluctuating hydrodynamics in the LBM. \cite{liu_simulation_2023, adhikari_fluctuating_2005}

Four primary techniques to model multicomponent or multiphase systems exist in the LBM literature. A quick review of the Shan-Chen or 
interparticle potential model  will be provided here from the perspective of a binary mixture. Additionally, the names and descriptions of 
other techniques will be described as well. The interparticle potential model or Shan-Chen model that adds a non-local density dependent force 
between two species, effectively modelling non-ideal mixing and recovering the Cahn-Hilliard equation. To alleviate the standard Shan-Chen implementations 
weaknesses of not being thermodynamically consistent and reducing the existence of spurious velocities,

In addition to the Shan-Chen model, the color gradient model, free energy model and interface tracking technique all 
allow for simulations of various types of multiphase and/or multicomponent flows. The color gradient model proposed by Rothman and Keller and 
implemented by Gunstensen et al relies upon modelling two particle distributions. These represent a binary fluid mixture with the collision step 
able to recover the hydrodynamics and non-ideal mixing dynamics. The free energy model utilizes phase field theory and constructs a free energy 
functional to recover interfacial dynamics and effects in a thermodynamically consistent manner. Often, a square gradient free energy is used for 
simplicity and ease of implementation. The mean field theory The non-ideal mixing dynamics are recovered during the "recoloring" step, of which the 
technique proposed by Latva-Kokko and Rothman is more commonly used for soft matter flows. \cite{liu_multiphase_2016}

>>>>

\section{Formation of particle stabilized emulsions}

When two partially miscible fluids are mixed, they phase separate into distinct phases to minimize the thermodynamic penalty associated with interfacial formation. 
To inhibit this phase separation, small-molecule surfactants are conventionally employed. A well-known example of such a material is soap, which facilitates the 
emulsification of dirt into droplets suspended in water through mechanical agitation. Surfactants function by reducing the surface tension between the dispersed 
phase (droplets) and the continuous phase (bulk fluid), thereby lowering the interfacial energy penalty associated with phase formation.

Without the addition of a stabilizer, the dispersed phase will coalescence to reduce the surface area of the interface. To facilitate the stability of small
droplets in the continuous phase, stabilizers such as surfactants, particles or bio-polymers are added. These additives reduce the interfacial energy between
phases characterized through the Pieranski model, $ G_{ads} = \sigma A_{rm} (1 - \cos{\theta_c})^2 $, where $G_{ads}$ represents the free energy reduction upon particle 
adsorption, $\sigma$ is the surface tension between the partially miscible fluids, and $ \theta_c $ is the contact angle of the particle. Surfactants and bio-polymers work
through reducing the surface tension between the dispersed and continuous phase. Particle stabilizers work to stabilize emulsions through reducing the interfacial area between
fluids.

@@

Compared to surfactant-stabilized emulsions, particle-stabilized emulsions exhibit greater resistance to coarsening, leading to renewed interest in their applications. 
The microstructural properties of these emulsions were extensively studied in the early 2000s by Binks and Lumsdon when they investigated the effect of particle size
on the resulting Pickering emulsion droplet size. \cite{binks_pickering_2001} Their findings 
indicated that the emulsion droplet radius follows the relationship $R_e \propto \frac{\phi_w}{\phi_p}$, where $\phi_w$ and $\phi_p$ represent the volume fractions of water and particles, 
respectively \cite{binks_pickering_2001}. Additionally, they identified several factors influencing the microstructure of Pickering emulsions, including fluid concentration 
and particle wettability. Neutrally wetting particles do not impose a preferential curvature on the interface, whereas non-neutrally wetting particles can lead to the 
formation of bridged droplets, capillary aggregates, and other non-spherical microstructures, shown in Figure \ref{fig:state_diagram_particle_emulsions}

\section{Bicontinuous interfacially jammed emulsion gels}

From Figure \ref{fig:state_diagram_particle_emulsions}, the formation of bijels is shown to require equal volume fraction of fluids and neutrally wetting particles. This
microstructure was first identified in 2005 when Stratford et al. used a multicomponent Lattice Boltzmann Method coupled with particles to characterize the microstructure
and dynamics of particles stabilized emulsions made with equal volume fractions of a binary fluid and suspended particles. \cite{stratford_colloidal_2005} They characterized
that upon initiation of the simulation, the fluids began to undergo spinodal decomposition. As the interface coarsens, particles adsorb onto it. When the particle surface area
and cross sectional area of the particles match, coarsening jams in place resulting in a tortuous co-continuous morphology. A visualization of this process is shown in 
Figure \ref{fig:bijel_coarsen}.

\begin{figure}
    \centering
    \includegraphics[scale = 0.3]{figures/introduction/bijel_coarsening.jpg}
    \caption{Initiation and arrest of spinodal decomposition as particles adsorb onto the interface, followed by jamming once the 
    interfacial area matches the cross sectional area of the adsorbed particles\cite{stratford_colloidal_2005}. 
    Reproduced from K. Stratford et al., Science 309, 2198-2201(2005) under license number 5966820525314.}
    \label{fig:bijel_coarsen}
\end{figure}

After their discovery in 2005, bijels were realized in experiments in 2007 using a 2-6 lutidine/water system stabilized by surface modified silica particles.
\cite{herzig_bicontinuous_2007} The 2-6 lutidine/water system has a lower critical solution temperature (LCST) of 34.1$^{\circ}C$ and a critical weight fraction of
approximately 28 \% lutidine. \cite{herzig_bicontinuous_2007} Herzig et al. began with a fluid mixture at the critical weight fraction and raised the 
temperature past the LCST, initiating phase separation. \cite{herzig_bicontinuous_2007} This process is known as Thermally Induced Phase Separation (TIPS). 
The domains of the fluid coarsened until they observed
gelation of the system. \cite{herzig_bicontinuous_2007} Since then, other systems based on small molecules, oligomers and polymers have been prepared. 
\cite{tavacoli_novel_2011, lee_bicontinuous_2010, bai_dynamics_2015, ching_rapid_2021} These works have been primarily based on
thermally induced phase separation although as this literature review will show modern bijel fabrication techniques have moved
away from this method to induce phase separation.

Computational simulations have investigated the effect of fluid density, surface tension and particle volume fraction.
\cite{jansen_bijels_2011} Jansen and Harting utilized a multicomponent Lattice Boltzmann method to assess the effect of fluid volume fraction,
particle volume fraction, surface tension, fluid density and particle contact angle on the microstructure of particle stabilized emulsions.
They demonstrated that the effect of the surface tension and initial fluid density were small
compared to the volume fraction of particles by measuring the length scale of the bijels. \cite{jansen_bijels_2011} The length of 
particles was proportional to the inverse of the particle volume fraction, $L \propto \frac{1}{\phi_p}$. \cite{jansen_bijels_2011} 
Jansen and Harting also illustrated that the emulsion state obtained, whether a Pickering emulsion or bijel, was most affected by 
the contact angle of the particle stabilizer and fluid ratio rather than the particle volume fraction. \cite{jansen_bijels_2011}

Further investigations of the bijel microstructure look at calculating the morphology of the bijel through topological, geometric 
or image analysis techniques. Topological techniques include the quantification of the genus and channel size distribution which
allow for determination of differences in the interfacial properties. \cite{chan_channel_2012} Chan and Thornton compared the topology
of bicontinuous systems obtained using the Cahn-Hilliard and Allen-Cahn equations and identified how they differed. \cite{chan_channel_2012}
Topological techniques have the advantage
of being able to characterize the microstructure of a bicontinuous material while also providing information about the connectivity of the
structure.

Geometric techniques such as
calculating the curvature of the bijel have also been used to identify the impact of particle size and quench rate on the microstructure
of bijels. Reeves et al. compared the Gaussian curvatures of bijels stabilized with micro and nanoparticles, as well as the effect of quench rate
on bijels formed using thermally induced phase separation to characterize that reduced particle size resulted in the interface maintaining a more negative
Gaussian curvature. \cite{reeves_particle-size_2015,reeves_quantitative_2016}
The interface of bijels are most similar to minimal surfaces, which indicates their mean curvature and Gaussian curvature are 0 and negative respectively. 
\cite{jinnai_interfacial_2001} Nanoparticles are able to conform to curved surfaces better as they are less disruptive to the curvature of the interface, resulting
in more effective stabilization of saddle like regions. \cite{reeves_quantitative_2016}

Image analysis techniques have also been used to characterize the microstructure of bijels, providing information on the connectivity, pore size
distribution, tortuosity and self similarity. \cite{mcdevitt_microstructural_2019,reeves_quantitative_2016} Connectivity can be derived from region
growing algorithms, labelling algorithms such as Hoshen-Kopelmann and segmentation. \cite{mcdevitt_microstructural_2019,reeves_quantitative_2016}
The pore size distribution can be obtained from watershed algorithm. \cite{mcdevitt_microstructural_2019}

TIPS has enabled batch synthesis of bijels in research laboratories. However, for bijels to become industrially viable, a continuous fabrication process is required.
This technique should also be flexible enough to produce tailored bijel microstructures to different applications. Solvent Transfer Induced Phase Separation (STrIPS) 
has been shown as a method to achieve this, shown in Figure \ref{fig:strips}. Figure \ref{fig:strips}a) shows the preparation of casting mixture in STrIPS. The casting
mixture is a ternary fluid system comprised of the fluids making up the bijel and a solvent. Figure \ref{fig:strips}b) demonstrates the removal of solvent from the
casting mixture, by flowing a co-solvent past the casting mixture, initiating phase separation. A narrow range of compositions can be used to ensure that the composition
of the phase separating mixture remains within the spinodal line of the demixed region of the phase diagram. Figure \ref{fig:strips} shows how varying the flow rate
of the casting mixture and the solvent removal medium is necessary to control the morphology of the bijel.

\begin{figure}[h]
    \centering
    \includegraphics[scale = 5]{figures/introduction/STrIPS.jpg}
    \caption{STrIPS in action. Extrusion of the bijel casting mixture into a non-solvent bath, followed by removal of solvent from 
             the casting mixture through diffusion. \cite{haase_continuous_2015}. Reproduced with permission from Wiley-VCH Verlag GmbH 
             \& Co. KGaA from \textit{Continuous Fabrication of Hierarchical and Asymmetric Bijel Microparticles, Fibers, and Membranes by 
             Solvent Transfer-Induced Phase Separation (STRIPS)}, Haase, M. F., Stebe, K. J., 
             \& Lee, D., \textit{Advanced Materials}, 27(44), 7065-7071, 2015}
    \label{fig:strips}
\end{figure}

STrIPS has been used extensively in a variety of systems ranging from bijels for battery electrode templates, to drug delivery medium, liquid in liquid printing and
intrinsically polymerizable bijels facilitating the production of free standing structures.
\cite{garcia_scalable_2019,thorson_bijel-templated_2019, amirfattahi_fabrication_2024, ching_rapid_2021} STrIPS is flexible to the casting mixture being used. The
only requirement a casting mixture used for STrIPS is that the phase diagram for the casting mixture must be well understood so that the solvent transfer step remains
within the spinodal at all times.

Another technique based on the removal of solvent is Vapor Induced Phase Separation(VIPS). \cite{wang_scalable_2020} A quarternary casting mixture
containing particles, solvent and two partially miscible liquids is prepared. The solvent and liquid species and compositions are carefully selected to ensure that 
removal of the solvent will cause phase separation of the two partially miscible liquids through the critical point of the phase diagram. This ensures that phase separation
only occurs through spinodal decomposition. One system that has been 
used is the water/hexanediol-diacrylate solvated by ethanol. Removal of ethanol at room temperature is sufficient to facilitate bijel formation, with this bijel
mixture composition being able to be spray or blade coated. This technique is limited by the temperature dependence of surface tension, meaning that the casting mixtures
that can be used are limited to solvents that can vaporize at room temperature.

Instead of relying upon spinodal decomposition to generate the bijel morphology, homogenization uses shear to join phase separating domains, resulting in a 
structure that has the properties of a bijel even if not fabricated from fluids undergoing spinodal decomposition. \cite{huang_bicontinuous_2017, cai_bijels_2017} 
This method is primarily used for nanoparticles under $50$ nm and it has been shown to work for multiple particle geometries. This method relies upon
shear to cause limited coalescence of droplets. As coarsening occurs, particles adsorb on the interface and when the interfacial area matches the area of the
adsorbed particles, the microstructure jams. Microstructure tuning of the method can be done through modifying the contact angle of the particles through addition of
surfactants or viscosity of the phase separating fluids through the addition of viscosity modified such as oligomers. Thus far, all these synthesis techniques use
spherical particles to stabilize the bijels that are generated.

\section{Ellipsoidal particles at interfaces}

Over the past decade, advancements in synthesis techniques have significantly expanded the ability to fabricate particles with anisotropic geometry and surface chemistry. 
The choice of synthesis method depends on the desired particle shape. For example, ellipsoidal particles can be readily produced through the mechanical deformation of 
polymer spheres, while dumbbell-shaped particles can be synthesized using microfluidic devices or emulsion templating. \cite{fei_magneto-capillary_2020} 
Square-shaped particles, on the other hand, can be 
generated through controlled crystallization \cite{morgan_understanding_2013}. The increasing variety of synthesis methods has reduced geometric constraints when exploring 
potential particle stabilizers for bijels \cite{wu_recent_2016}.

@@

Anisotropic particles exhibit shape-dependent properties due to their ability to induce multipolar interactions by deforming the interface to maintain a mean curvature of 
zero, thereby satisfying the Young-Laplace equation, as illustrated in Figure \ref{fig:anisotropic_particle_interface} \cite{loudet_capillary_2005, cheng_shape-anisotropic_2013}.
This property has been leveraged to facilitate directed assembly and migration through modified capillary forces imparted through curvature gradients at interfaces.
\cite{cavallaro_curvature-driven_2011, read_dimerization_2020, sharifi-mood_curvature_2015}. Simulations have shown that ellipsoidal particles can adopt metastable orientations depending upon the perturbations 
the particle experiences. \cite{gunther_lattice_2013}
% The adsorption process is affected by the shape of particles used as they pack differently onto the interface \cite{hijnen_bijels_2015, daware_emulsions_2015,carmack_diverse_2017}. It has also been suggested that adsorption dynamics of ellipsoidal particles at liquid interfaces are driven completely by viscous forces even if the timescales of adsorption are driven by particle properties \cite{Coertjens2017}. Some guiding equations to calculate the interfacial area of an ellipsoidal particle are provided \cite{gunther_timescales_2014, Davies2014}.

Ellipsoidal particles have been used as stabilizers in bijels, resulting in bijels that have smaller domains than those made with spherical particles
at the same volume fraction, identified to arise from the  greater surface area to volume ratio of ellipsoidal particles compared to spherical particles. 
\cite{gunther_timescales_2014} The evolution of the microstructure is also distinct to bijels stabilized with spheres, demonstrating two timescales at
early and later timesteps. \cite{gunther_timescales_2014} The short timescale is during formation of the bijel, when particles are adsorbed onto the
interface and rotate to align with the interface. At longer timescales, capillary interactions between particles lead to reordering of particles
at the interface, leading to a different interfacial configuration and microstructure. \cite{gunther_timescales_2014}

Rod like particles have also been used as stabilizers in experiments, identifying that the length scale of the pores of 
the bijels follows $ L \propto \frac{1}{\phi_p} $ even if the proportionality constant is different. While the proportionality constant depends on 
particle geometry, this trend has been observed across various particle shapes 
\cite{hijnen_bijels_2015, madivala_exploiting_2009, gunther_timescales_2014, daware_emulsions_2015, loudet_capillary_2005, cheng_shape-anisotropic_2013}.
The dynamics of bijels stabilized by ellipsoidal particles are also distinct from bijels stabilized by spheres, caused by reorientation of particle
stabilizers at the interface.

\section{Magnetic fields applied to particles at interfaces}

Magnetic fields offer a non-invasive and targeted method for manipulating the behavior of particles. Magnetic particles can be fabricated from
coatings applied onto particles through deposition techniques or a core-shell particle that encapsulates a magnetic core with a non-magnetic shell. 
\cite{fei_magneto-capillary_2020, nakayama_stimuli-responsive_2018} These effects arise due to the interplay between the magnetic torque or force 
acting on the particles and the capillary interactions that govern their behavior at the interface. One theory developed for neutrally wetting ellipsoidal 
particles is Bresme-Faraudo theory. \cite{bresme_orientational_2007, davies_interface_2014}

A wide range of behaviors has been observed when magnetic fields are applied to interfacial particles. Individual particles tend to align their magnetic 
dipole moments along the direction of the field, causing them to reorient relative to the interface. \cite{kim_bijels_2010}
For anisotropic particles such as ellipsoids, this results in a tilt or rotation that changes their interfacial footprint and local curvature. 
\cite{davies_assembling_2014, davies_interface_2014} The orientation of the particle is a function of the capillary interactions with the interface that 
prefers to keep the particle flat on the interface and the magnetic field that is attempting to rotate the particle out of the interface. The magnetic force
experienced by the particle is a function of the applied field strength, magnetic susceptibility, size of the particle, shape of the particles 
and aspect ratio of the particle.

The equilibrium and metastable orientations of non-spherical particles at fluid interfaces are governed by a complex interplay of particle shape, 
interfacial tension, and external forces such as magnetic fields. Depending upon the surface wettability and aspect ratio of the particle, non-spherical
particles have been observed to adopt several stable and metastable orientations. \cite{morgan_understanding_2013, newton_influence_2014}

Studies on superellipsoidal hematite particles have demonstrated that isolated 
anisotropic particles at the oil–water interface can adopt multiple stable and metastable orientations depending on their aspect ratio and surface 
wettability. Specifically, particles exhibit orientations where their major or minor axes align either parallel or perpendicular to the interface, 
with each configuration corresponding to a local or global energy minimum depending on the interfacial energy landscape \cite{1}. Lattice Boltzmann 
simulations further support these findings, revealing that particle aspect ratio significantly influences the contact angle and immersion depth, 
which in turn affect the energetic favorability of different orientations \cite{3}. In the presence of a magnetic field, these equilibrium states can 
shift due to the torque exerted on the particles, aligning their magnetic moments with the field direction. Experimental observations confirm that 
magnetic anisotropic particles, such as hematite ellipsoids, can be reoriented at the interface by applying a magnetic field, leading to controlled 
transitions between metastable and stable states \cite{2}. These studies collectively show that the equilibrium orientation of anisotropic particles 
is not singular, but rather a function of particle geometry, field strength, and interfacial anchoring, with metastable states playing a significant 
role in the dynamic reconfiguration of interfacial structures.

In many particle system, the magnetic field driven interfacial reorientation lead to the formation of linear chains or clusters of particles on flat interfaces. 
\cite{davies_assembling_2014, davies_interface_2014,newton_influence_2014, newton_capillary_2018} The orientation of particles relative to one another is also
a function of the number of particles in contact and the applied field strength owing to the magnetic field driven interactions between each particle. 
\cite{newton_influence_2014, newton_capillary_2018} In Pickering emulsions, this has enabled field-controlled droplet deformation, coalescence, 
and motility, illustrating the potential for on-demand manipulation of interfacial structures. \cite{melle_pickering_2005, tham_magnetophoresis_2021}

\textcolor{blue}{stuff about how magnetic fields affect rheology in pickering emulsions}

\section{Rheological models and shear response of particle stabilized emulsions}

@@

Ellipsoidal particles influence the rheological properties and stability of Pickering emulsions. When investigating the effect of rod-like particles on the
rheology of emulsion droplets, Madivala et al. showed that droplets stability under shear increased as the aspect ratio of the rod-like particle increased.
\cite{madivala_exploiting_2009} This was attributed to the more effective interfacial coverage of particles and interface deformations inducing attractive
capillary interactions between particles, reinforcing the particle monolayer to applied shear. \cite{madivala_exploiting_2009} Investigations into the rheology of
Pickering droplets stabilized by graphene nanosheets showed that the dicsc like particles had posessed increased viscosity and viscoelasticity compared to the same 
system synthesized with spherical particles. \cite{imperiali_simple_2014} The elasticity of the interface has been shown to be a function of the particle stabilizer
used. \cite{sun_assembly_2013}

Magnetically responsive Pickering emulsions exhibit rheological behavior that can be tuned by external magnetic fields, making them attractive for applications 
requiring on-demand flow control and structural reconfiguration. \cite{qiao_magnetorheological_2012, melle_pickering_2005} Upon application of a field, particles
order at the interface. This tunable rheology enables reversible transitions between fluid and gel-like states, which is particularly valuable in 
applications such as enhanced oil recovery, drug delivery, and soft robotics \cite{tham_magnetophoresis_2021}.

@@

\section{Stimuli response in bijels}

@@
